---
title: "Chapter 16 - Hoan"
output: github_document
date: "2023-02-19"
---

1.  Introduction

Not every years has 365 days -\> a leap year? -\> daylight saving times (DST) ( 23 hours or 25 hours/day)

1.1. Load package

```{r}
library(tidyverse)
library(lubridate)
library(nycflights13)
```

1.2. Creating date/times

-   **A date : Tibbles print this as \<date\>**

-   A time : <time>

-   **A date-time: \<dttm\> -** complicated because of need to handle time zones

To get the current date/date-time

```{r}
today()
now()
```

There are three ways you're likely to create a date/time:

-   From a string

-   From individual date-time components

-   From an existing date/time object

1.2.1. From strings

-   **lubridate \>** specitfy the order: year, month, day \> arrange "y", "m", "d" in the same order

-   **ymd mdy dmy**

```{r}
ymd("2017-01-31")
mdy("January 31st, 2017")
dmy("31-Jan-2017")
```

-   Unquoted numbers : ymd()

```{r}
ymd(20170131)
```

-   Create a date time: add an underscore and one or more of "h"/"m"/"s" (hours/minute/second)

    ```{r}
    ymd_hms("2017-01-31 20:11:59")
    mdy_hm("01/31/2017 08:01")

    ```

-   Supply a timezone

```{r}
ymd(20170131, tz = "UTC")

```

1.2.2. From individual components

-   Review the flights data \> select: year, month, day, hour, minute

    ```{r}
    flights %>% 
      select(year, month, day, hour, minute)
    ```

-   Create a date/time : make_date () for dates, make_datetime() for date-times

    ```{r}
    flights %>%
      select(year, month, day, hour, minute) %>% 
      mutate(departure = make_datetime(year, month, day, hour, minute))
    ```

-   modulus arithmetic to pull out the hour and minute components

```{r}
make_datetime_100 <- function(year, month, day, time) {
  make_datetime(year, month, day, time %/% 100, time %% 100)
}

flights_dt <- flights %>% 
  filter(!is.na(dep_time), !is.na(arr_time)) %>% 
  mutate(
    dep_time = make_datetime_100(year, month, day, dep_time),
    arr_time = make_datetime_100(year, month, day, arr_time),
    sched_dep_time = make_datetime_100(year, month, day, sched_dep_time),
    sched_arr_time = make_datetime_100(year, month, day, sched_arr_time)
  ) %>% 
  select(origin, dest, ends_with("delay"), ends_with("time"))

flights_dt
```

-   Visualise the distribution of departure times

    ```{r}
    flights_dt %>% 
      ggplot(aes(dep_time)) + 
      geom_freqpoly(binwidth = 86400) # 86400 seconds = 1 d

    ```

-   Or within a single day

    ```{r}
    flights_dt %>% 
      filter(dep_time < ymd(20130102)) %>% 
      ggplot(aes(dep_time)) + 
      geom_freqpoly(binwidth = 600) # 600 s = 10 minutes
    ```

1.2.3. From other types

-   Switch a date-time /a date : as_datetime()

    ```{r}
    as_datetime(today())
    as_date(now())

    ```

-   date/times: numeric offsets from the Unix Epoch

```{r}
as_datetime(60 * 60 * 10)
as_date(365 * 10 + 2)
```

1.3. Date-time components

1.3.1. Getting components

-   year(), month (), mday, yday, wday, hour, minute, second

```{r}
datetime <- ymd_hms("2023-02-20 20:59:21")

year(datetime)
month(datetime)
mday(datetime)
yday(datetime)
wday(datetime)
```

-   month, wday: label = True \> the abbreviated name of the month or day of the week; abbr = FALSE to return the full name

    ```{r}
    month(datetime, label = TRUE)
    wday(datetime, label = TRUE, abbr = FALSE)
    ```

-   wday () : see that more flights depart during the weel than on the weekend

    ```{r}
    flights_dt %>% 
      mutate(wday = wday(dep_time, label = TRUE)) %>% 
      ggplot(aes(x = wday)) +
        geom_bar()
    ```

-   look at the average departure

    ```{r}
    flights_dt %>% 
      mutate(minute = minute(dep_time)) %>% 
      group_by(minute) %>% 
      summarise(
        avg_delay = mean(arr_delay, na.rm = TRUE),
        n = n()) %>% 
      ggplot(aes(minute, avg_delay)) +
        geom_line()
    ```

-   look at the scheduled departure

```{r}
sched_dep <- flights_dt %>% 
  mutate(minute = minute(sched_dep_time)) %>% 
  group_by(minute) %>% 
  summarise(
    avg_delay = mean(arr_delay, na.rm = TRUE),
    n = n())

ggplot(sched_dep, aes(minute, avg_delay)) +
  geom_line()
```

-   "nice" departutre times -\> always be alert

```{r}
ggplot(sched_dep, aes(minute, n)) +
  geom_line()
```

1.3.2. Rounding

-   Roung the date to a nearby unit of time: floor_date (round down, roun_date, ceiling_date (round_up)

```{r}
flights_dt %>% 
  count(week = floor_date(dep_time, "week")) %>% 
  ggplot(aes(week, n)) +
    geom_line()
```

1.3.3. Setting components

-   use function to set components of a date /time

```{r}
(datetime <- ymd_hms("2016-07-20 12:34:56"))
year(datetime) <- 2023
datetime
month(datetime) <- 02
datetime
hour(datetime) <- hour(datetime) + 1
datetime

```

-   No modify -\> use update()

    ```{r}
    update(datetime, year = 2024, month = 2, mday = 6, hour = 8)

    ```

-   values too big -\> roll-over

```{r}
ymd("2023-01-01") %>% 
  update(mday = 4)

ymd("2023-01-01") %>% 
  update(hour = 24 )

```

-   update()-\> distribution of flights across the course of the day for every day of the year

```{r}
flights_dt %>% 
  mutate(dep_hour = update(dep_time, yday = 1)) %>% 
  ggplot(aes(dep_hour)) +
    geom_freqpoly(binwidth = 300)
```

1.4. Time spans

-   3 classes: durations/periods/intervals

    1.4.1. Durations

```{r}
# How old is Hadley?
h_age <- today() - ymd(19791014)
h_age

```

```{r}
as.duration(h_age)
```

```{r}
dseconds(15)
dminutes(10)
dhours(c(12, 24))
ddays(0:5)
dweeks(3)
dyears(1)
```

-   add and multiply durations

```{r}
2 * dyears(1)
dyears(1) + dweeks(12) + dhours(15)
```

-   add and substract durations to and from days

```{r}
tomorrow <- today() + ddays(1)
last_year <- today() - dyears(1)
tomorrow
last_year
```

-   Get an unexpected results

    ```{r}
    one_pm <- ymd_hms("2016-03-12 13:00:00", tz = "America/New_York")

    one_pm
    one_pm + ddays(1)
    ```

1.4.2. Periods

-   work with "human" times

    ```{r}
    one_pm
    one_pm + days(1)
    ```

```{r}
seconds(15)
minutes(10)
hours(c(12, 24))
days(7)
months(1:6)
weeks(3)
years(1)
```

-   Addd/multiply periods

```{r}
10 * (months(6) + days(1))
days(50) + hours(25) + minutes(2)
```

```{r}
# A leap year
ymd("2016-01-01") + dyears(1)
ymd("2016-01-01") + years(1)

one_pm + ddays(1)
one_pm + days(1)

```

-   fix an oddity related to flight dates

```{r}
flights_dt %>% 
  filter(arr_time < dep_time) 
```

-   fix by adding days(1) to the arrival time of each overnight flight

```{r}
flights_dt <- flights_dt %>% 
  mutate(
    overnight = arr_time < dep_time,
    arr_time = arr_time + days(overnight * 1),
    sched_arr_time = sched_arr_time + days(overnight * 1)
  )

```

```{r}
flights_dt %>% 
  filter(overnight, arr_time < dep_time) 
```

1.4.3. Intervals

```{r}
years(1) / days(1)

```

```{r}
next_year <- today() + years(1)
(today() %--% next_year) / ddays(1)

```
