---
title: "Iteration"
output: github_document
date: "2023-03-16"
---

```{r setup, include=FALSE}
install.packages("tidyverse")
install.packages("purrr")
library(tidyverse)
library(purrr)
```

1.  **For loops**

```{r}
df <- tibble(
a = rnorm(10),
b = rnorm(10),
c = rnorm(10),
d = rnorm(10)
)
df
```

Tính trung vị cho từng cột :

```{r}
median(df$a)
median(df$b)
median(df$c)
median(df$d)
```

Tạo 1 vòng lặp để tính trung vị :

```{r}
output <- vector("double", ncol(df))  # 1. output
for (i in seq_along(df)) {            # 2. sequence
  output[[i]] <- median(df[[i]])      # 3. body
}
output
```

```{r}
y <- vector("double", 0)
seq_along(y) 
1:length(y)
```

**Mỗi vòng lặp for có 3 thành phần :**

-Output

vector("type",length) : create an empty vector of given length

"type" : "logical", "integer", "double", "character", etc

-Sequence : for ... in ...

-Body : function, gán biến, etc...

**Syntax:**

output \<- ....

for (i in ...) {

function

}

output

1.  Compute the mean of every column in `mtcars`.

```{r}
mtcars

output <- vector("double",ncol(mtcars))
for (i in seq_along(mtcars)) {
  output[[i]] <- mean(mtcars[[i]])
  
}
output
```

Determine the type of each column in nycflights13::flights.

```{r}
nycflights13::flights
output <- vector("character",ncol(nycflights13::flights))
for (i in seq_along(nycflights13::flights)) {
  output[[i]] <- typeof(nycflights13::flights[[i]])
}
output
```

## **2. For loop variations**

There are four variations on the basic theme of the for loop:

1.  Modifying an existing object, instead of creating a new object.

    ```{r}
    df <- tibble(
      a = rnorm(10),
      b = rnorm(10),
      c = rnorm(10),
      d = rnorm(10)
    )
    rescale01 <- function(x) {
      rng <- range(x, na.rm = TRUE)
      (x - rng[1]) / (rng[2] - rng[1])
    }

    df$a <- rescale01(df$a)
    df$b <- rescale01(df$b)
    df$c <- rescale01(df$c)
    df$d <- rescale01(df$d)
    ```

1.Output: đã có

2.Sequence : DF = list of columns --\> seq_along(df)

3.Body : function rescale01()

```{r}
for (i in seq_along(df)) {
  df[[i]] <- rescale01(df[[i]])
}
```

[[ i ]] : take a single element

2.  Looping patterns (names, values, indices)

    3 basic ways:

    -loop over the numeric indices: for (i in seq_along(xs)) -loop over the elements: for (x in xs)

    -loop over the names : for (nm in names(xs))

```{r}
results <- vector("list", length(x))
names(results) <- names(x)
```

```{r}
for (i in seq_along(x)) {
  name <- names(x)[[i]]
  value <- x[[i]]
}
```

3.  Handling outputs of unknown length.

```{r}
means <- c(0, 1, 2)

output <- double()
for (i in seq_along(means)) {
  n <- sample(100, 1)
  output <- c(output, rnorm(n, means[[i]]))
}
str(output)
```

```{r}
out <- vector("list", length(means))
for (i in seq_along(means)) {
  n <- sample(100, 1)
  out[[i]] <- rnorm(n, means[[i]])
}
str(out)

str(unlist(out))
```

purrr::flatten_dbl() : flatten a list of vectors into a single vector

4.  Handling sequences of unknown length.

while (condition) { #body }

2 thanh phan : condition, body

Co the viet lai vong lap for duoi dang vong lap while nhung khong the lam nguoc lai

for (i in seq_along(x)) { \# body }

i \<- 1 while (i \<= length(x)) { \# body i \<- i + 1 }

```{r}
flip <- function() sample(c("T", "H"), 1)

flips <- 0
nheads <- 0

while (nheads < 3) {
  if (flip() == "H") {
    nheads <- nheads + 1
  } else {
    nheads <- 0
  }
  flips <- flips + 1
}
flips
```

3.  **For loops vs. functionals**

```{r}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
```

```{r}
output <- vector("double", length(df))
for (i in seq_along(df)) {
  output[[i]] <- mean(df[[i]])
}
output
```

```{r}
col_mean <- function(df) {
  output <- vector("double", length(df))
  for (i in seq_along(df)) {
    output[i] <- mean(df[[i]])
  }
  output
}
```

```{r}
col_median <- function(df) {
  output <- vector("double", length(df))
  for (i in seq_along(df)) {
    output[i] <- median(df[[i]])
  }
  output
}
col_sd <- function(df) {
  output <- vector("double", length(df))
  for (i in seq_along(df)) {
    output[i] <- sd(df[[i]])
  }
  output
}
```

f1 \<- function(x) abs(x - mean(x)) \^ 1 f2 \<- function(x) abs(x - mean(x)) \^ 2 f3 \<- function(x) abs(x - mean(x)) \^ 3

f \<- function(x, i) abs(x - mean(x)) \^ i

```{r}
col_summary <- function(df, fun) {
  out <- vector("double", length(df))
  for (i in seq_along(df)) {
    out[i] <- fun(df[[i]])
  }
  out
}
col_summary(df, median)
col_summary(df, mean)
```

4.  **The map functions :**

    loi ich : de viet, de doc hon vong for (ko lien quan van de toc do)

map() makes a list. map_lgl() makes a logical vector. map_int() makes an integer vector. map_dbl() makes a double vector. map_chr() makes a character vector.

```{r, error=TRUE}
map_dbl(df, mean)

map_dbl(df, median)

map_dbl(df, sd)

```

```{r, error=TRUE}
df %>% map_dbl(mean)

df %>% map_dbl(median)

df %>% map_dbl(sd)

```

differences between map\_\*() and col_summary() : -All purrr functions are implemented in C. This makes them a little faster at the expense of readability.

-The second argument, .f, the function to apply, can be a formula, a character vector, or an integer vector. You'll learn about those handy shortcuts in the next section.

-map\_\*() uses ... ([dot dot dot]) to pass along additional arguments to .f each time it's called:

```{r}
map_dbl(df, mean, trim = 0.5)

```

```{r}
z <- list(x = 1:3, y = 4:5)
map_int(z, length)

```

Shortcuts

```{r}
models <- mtcars %>% 
  split(.$cyl) %>% 
  map(function(df) lm(mpg ~ wt, data = df))
models
```

```{r}
models <- mtcars %>% 
  split(.$cyl) %>% 
  map(~lm(mpg ~ wt, data = .))
models
```

```{r}
models %>% 
  map(summary) %>% 
  map_dbl(~.$r.squared)
```

```{r}
models %>% 
  map(summary) %>% 
  map_dbl("r.squared")
```

```{r}
x <- list(list(1, 2, 3), list(4, 5, 6), list(7, 8, 9))
x
x %>% map_dbl(2)
```

Base R

-lapply() : giống với map(), ngoại trừ map() phù hợp với tất cả các chức năng khác trong purrr và có thể sử dụng phím tắt cho .f. -sapply() : bao boc xung quanh lappy(), tu dong don gian hoa dau ra nhung ko bao gio biet minh se nhan dc output gi

```{r}
x1 <- list(
  c(0.27, 0.37, 0.57, 0.91, 0.20),
  c(0.90, 0.94, 0.66, 0.63, 0.06), 
  c(0.21, 0.18, 0.69, 0.38, 0.77)
)
x2 <- list(
  c(0.50, 0.72, 0.99, 0.38, 0.78), 
  c(0.93, 0.21, 0.65, 0.13, 0.27), 
  c(0.39, 0.01, 0.38, 0.87, 0.34)
)

threshold <- function(x, cutoff = 0.8) x[x > cutoff]
x1 %>% sapply(threshold) %>% str()

x2 %>% sapply(threshold) %>% str()

```

-vapplpy(df, is.numeric, logical) : giai phap an toan cho sapply(),cung cap doi so xac dinh type, co the tao ra cac matrice (map() chi tao cac vector)

5.  **Dealing with failure**

    safely() : nhan vao 1 function va tra ve 1 phien ban da sua doi, list cua 2 phan tu: -ket qua ban dau, neu co loi se la NULL -error object, neu operation thanh cong thi se la NULL.n

try() : thinh thoang tra ve ket qua ban dau thinh thoang tra ve error object

```{r}
safe_log <- safely(log)
str(safe_log(10))

str(safe_log("a"))

```

```{r}
x <- list(1, 10, "a")
y <- x %>% map(safely(log))
str(y)

```

purrr::transpose(): tra ve 2 lists : all errors va all outputs

```{r}
y <- y %>% transpose()
str(y)

```

```{r}



y$result[is_ok] %>% flatten_dbl()

```

possibly() : giong ham safely() nhung don gian hon (dua vao gia tri mac dinh)

```{r}
x <- list(1, 10, "a")
x %>% map_dbl(possibly(log, NA_real_))
```

quietly() : giong ham safely() nhung thay tra ve error, no tra ve output, messages va warning

6.  **Mapping over multiple arguments**

map2() , pmap() : iterate multiple related inputs in parallel pmap() : takes a list of arguments

```{r}
mu <- list(5, 10, -3)
mu %>% 
  map(rnorm, n = 5) %>% 
  str()

```

Thay doi do lech chuan

```{r}
sigma <- list(1, 5, 10)
seq_along(mu) %>% 
  map(~rnorm(5, mu[[.]], sigma[[.]])) %>% 
  str()

```

```{r}
map2(mu, sigma, rnorm, n = 5) %>% str()

```

Note that the arguments that vary for each call come before the function; arguments that are the same for every call come after.

```{r}
n <- list(1, 3, 5)
args1 <- list(n, mu, sigma)
args1 %>%
  pmap(rnorm) %>% 
  str()

```

If you don't name the list's elements, pmap() will use positional matching when calling the function.

```{r}
args2 <- list(mean = mu, sd = sigma, n = n)
args2 %>% 
  pmap(rnorm) %>% 
  str()
```

```{r}
params <- tribble(
  ~mean, ~sd, ~n,
    5,     1,  1,
   10,     5,  3,
   -3,    10,  5
)
params %>% 
  pmap(rnorm)
```

Invoking different functions

```{r}
f <- c("runif", "rnorm", "rpois")
param <- list(
  list(min = -1, max = 1), 
  list(sd = 5), 
  list(lambda = 10)
)
```

invoke_map() : list of function, list of giving the arguments, n

```{r}
invoke_map(f, param, n = 5) %>% str()

```

```{r}
sim <- tribble(
  ~f,      ~params,
  "runif", list(min = -1, max = 1),
  "rnorm", list(sd = 5),
  "rpois", list(lambda = 10)
)
sim %>% 
  mutate(sim = invoke_map(f, params, n = 10))
```

8.  **Walk** :

    giong ham map(), muon goi 1 ham bi tac dung phu cua no thay vi gia tri tra ve cua no Thuong dung de export output ra man hinh hoac save ra dia

```{r}
x <- list(1, "a", 3)

x %>% 
  walk(print)

```

walk2() , pwalk() : save each file to the corresponding location on disk if we have a list of plots and a vector of file names

```{r}
library(ggplot2)
plots <- mtcars %>% 
  split(.$cyl) %>% 
  map(~ggplot(., aes(mpg, wt)) + geom_point())
paths <- stringr::str_c(names(plots), ".pdf")

pwalk(list(paths, plots), ggsave, path = tempdir())
```

9.  **Other patterns of for loops**

Predicate functions :return a single TRUE or FALSE keep(), discard() :

```{r}
iris %>% 
  keep(is.factor) %>% 
  str()

iris %>% 
  discard(is.factor) %>% 
  str()

```

some() and every() determine if the predicate is true for any or for all of the elements.

```{r}
x <- list(1:5, letters, list(10))

x %>% 
  some(is_character)

x %>% 
  every(is_vector)
```

detect() finds the first element where the predicate is true; detect_index() returns its position.

```{r}
x <- sample(10)
x

x %>% 
  detect(~ . > 5)

x %>% 
  detect_index(~ . > 5)
```

head_while() and tail_while() take elements from the start or end of a vector while a predicate is true:

```{r}
x %>% 
  head_while(~ . > 5)

x %>% 
  tail_while(~ . > 5)
```

Reduce and accumulate

```{r, error=TRUE}
#install.packages("purrr")
#library(dplyr)
#library(purrr)
#packageVersion("purrr")
#unloadNamespace("purrr")

dfs <- list(
  age = tibble(name = "John", age = 30),
  sex = tibble(name = c("John", "Mary"), sex = c("M", "F")),
  trt = tibble(name = "Mary", treatment = "A")
)

dfs %>% reduce(full_join)

```

```{r, error=TRUE}
vs <- list(
  c(1, 3, 5, 6, 10),
  c(1, 2, 3, 7, 8, 10),
  c(1, 2, 3, 4, 8, 9, 10)
)

vs %>% reduce(intersect)
```

reduce() takes a "binary" function (i.e. a function with two primary inputs), and applies it repeatedly to a list until there is only a single element left.

accumulate() is similar but it keeps all the interim results.

```{r, error=TRUE}
x <- sample(10)
x
x %>% accumulate(`+`)
```
